Simple repression draft

### 2020-06-02 doc/section_05_discussion.tex
RP June 2
Intro has references to Muller-Hill
Discussion expanded
bib has many new entries

With the results of the constitutive promoter in hand, we then fix the
parameters associated with this class of promoters and use them as input for
evaluating the noise in gene expression for the simple repression motif itself.
This allows us to provide a single overarching analysis of both the constitutive
and simple repression architectures using one simple model and corresponding set
of self-consistent parameters, demonstrating not only a predictive framework,
but also reconciling the equilibrium and non-equilibrium views of the same
simple repression constructs. More specifically, we obtained values for the
transcription factor association and dissociation rates by performing Bayesian
inference on the ful mRNA distribution for data obtained from simple-repression
promoters with varying number of transcription factors per cell and affinity of
such transcription factors for the binding site. The free energy value obtained
from these kinetic rates -- computed as the log ratio of the rates -- agrees
with previous inferences performed only from mean gene expression measurements,
that assumed an equilibrium rather than a kinetic
framework~\cite{Garcia2011a,Razo-Mejia2018}.

@@

\mrm{We need to decide ASAP if this is going to be included in the discussion or not. My vote is no, but Jupiter must decide.}
\mmnote{Do all my approximations work simply because galK happens to have a low
duty cycle of supercoiling, $\beta/\alpha$ in Chong2014 picture? A potentially
interesting test: put in the dozen or so Hernan constructs at some other locus,
preferably one of the ones Chong identified as having a large duty cycle (large
$\beta/\alpha$), i.e., the opposite limit of galK. Then just measure
fold-changes. You don't even need to do FISH, although that would be even
better. If their model is right, $\rho$ should be approximately set by the duty
cycle of supercoiling, NOT by RNAP kinetics.
\textit{This is a NICE prediction for future work.} \textit{cspE}
is claimed nonessential, prediction is you drop HG reporter constructs in there
and the inferred binding energies of all lacI operators shift the same amount,
by about $\ln(3)\sim 1kT$. Not huge, but detectable, although might be
confounding effects from moving to a new locus. With FISH measurements, $k^+$
and $k^-$ in Fig 2, model 4 should now be separately inferrable if Chong2014
measurements of $\beta/\alpha$ are correct. This would require an obnoxious
amount of cloning, but once you did that, the measurements would be same old
stuff that RP lab does.}

A direct \mmnote{if cloning intensive} future experiment would be
to design a payload with a promoter driving a FISH reporter,
integrate that payload at a variety of loci scattered around the genome,
and carry out FISH measurements of the resulting strains.
\mmnote{
Did Brewster already do this? I remember he integrated arrays of
competitor binding sites all over the genome, but did he put
lacUV5 + reporter scattered around?}
Obviously integration loci would need to be carefully chosen so
that one could plausibly hope the deletion and integration did
not disturb the cell's physiology too badly.

\mrm{Rather than proposing more experiments, I feel that we are missing a closing paragraph that puts all these findings in context. Where do we go from here as a community? Things such as the next building block, i.e. simple activation, and then from there scale up the complexity until we can get a full gene regulatory network kind of thing. Rather than going small into tiny details specific to this, let's step back and look at the big picture.}

>>>>

With the results of the constitutive promoter in hand, we then fix the
parameters associated with this class of promoters and use them as input for
evaluating the noise in gene expression for the simple repression motif itself.
This allows us to provide a single overarching analysis of both the constitutive
and simple repression architectures using one simple model and corresponding set
of self-consistent parameters, demonstrating not only a predictive framework,
but also reconciling the equilibrium and non-equilibrium views of the same
simple repression constructs. More specifically, we obtained values for the
transcription factor association and dissociation rates by performing Bayesian
inference on the full mRNA distribution for data obtained from simple-repression
promoters with varying number of transcription factors per cell and affinity of
such transcription factors for the binding site. The free energy value obtained
from these kinetic rates -- computed as the log ratio of the rates -- agrees
with previous inferences performed only from mean gene expression measurements,
that assumed an equilibrium rather than a kinetic
framework~\cite{Garcia2011a,Razo-Mejia2018}.

@@

\textcolor{red}{RP: I relegated all remaining comments to the background.  Muir
can look at them to see if he really wants them.  I agreed with Manuel that the stuff
below should not be included in the Discussion.  ONE THING FOR SURE - Muller-Hill
and Oehler need to get a lot of love in references at beginning.}
In summary, this work took up the challenge of exploring the extent
to which a single specific mechanistic model of the simple-repression regulatory
architecture suffices to explain the broad sweep of experimental
data for this system.  Pioneering early experimental efforts from
the M\"{u}ller-Hill lab established the simple-repression motif
as an arena for the quantitative dissection of regulatory response in
bacteria, with similar beautiful work emerging in examples such as the
arabinose and gal operons as well~\cite{Dunn1984b, Oehler1990, Oehler1994, Schleif2000,Weickert1993, Semsey2002, SwintKruse2009}.  In light of a new generation of precision measurements
on these systems, the definition of what it means to understand them can now
be formulated as a rigorous quantitative question, and many aspects of
the simple repression motif seem to merit the claim that they are sufficiently
well understood to use that understanding to design new versions of that
architecture based upon predictions about how repressor copy number and DNA binding site strength
control expression.  In our view, the next step in the progression is
to first perform similar rigorous analyses of the fundamental ``basis set'' of regulatory
architectures such as the simple activation motif that is regulated by
a single activator binding site, the repressor-activator architecture, mediated by
the binding of both a single activator and a single repressor and beyond.
With the individual input-output functions in hand, similar quantitative
dissections including the rigorous analysis of their tuning parameters can be undertaken
for the ``basis set'' of full gene-regulatory networks such as switches, feed-forward architectures and
oscillators for example, building upon the recent impressive bonanza of efforts from
systems biologists and synthetic biologists~\cite{Milo2002,Alon2007}.

%\mrm{We need to decide ASAP if this is going to be included in the discussion or not. My vote is no, but Jupiter must decide.}
%\mmnote{Do all my approximations work simply because galK happens to have a low
%duty cycle of supercoiling, $\beta/\alpha$ in Chong2014 picture? A potentially
%interesting test: put in the dozen or so Hernan constructs at some other locus,
%preferably one of the ones Chong identified as having a large duty cycle (large
%$\beta/\alpha$), i.e., the opposite limit of galK. Then just measure
%fold-changes. You don't even need to do FISH, although that would be even
%better. If their model is right, $\rho$ should be approximately set by the duty
%cycle of supercoiling, NOT by RNAP kinetics.
%\textit{This is a NICE prediction for future work.} \textit{cspE}
%is claimed nonessential, prediction is you drop HG reporter constructs in there
%and the inferred binding energies of all lacI operators shift the same amount,
%by about $\ln(3)\sim 1kT$. Not huge, but detectable, although might be
%confounding effects from moving to a new locus. With FISH measurements, $k^+$
%and $k^-$ in Fig 2, model 4 should now be separately inferrable if Chong2014
%measurements of $\beta/\alpha$ are correct. This would require an obnoxious
%amount of cloning, but once you did that, the measurements would be same old
%stuff that RP lab does.}
%
%A direct \mmnote{if cloning intensive} future experiment would be
%to design a payload with a promoter driving a FISH reporter,
%integrate that payload at a variety of loci scattered around the genome,
%and carry out FISH measurements of the resulting strains.
%\mmnote{
%Did Brewster already do this? I remember he integrated arrays of
%competitor binding sites all over the genome, but did he put
%lacUV5 + reporter scattered around?}
%Obviously integration loci would need to be carefully chosen so
%that one could plausibly hope the deletion and integration did
%not disturb the cell's physiology too badly.
%
%\mrm{Rather than proposing more experiments, I feel that we are missing a closing paragraph that puts all these findings in context. Where do we go from here as a community? Things such as the next building block, i.e. simple activation, and then from there scale up the complexity until we can get a full gene regulatory network kind of thing. Rather than going small into tiny details specific to this, let's step back and look at the big picture.}
